Version A consistency is an important aspect of all software projects. As you should have learned by now, nothing is set in stone in the software world. Software evolves and changes over time. Such changes often need to be acknowledged in advance, either by running a series of tests against the new version or by making changes to the consuming code itself. Ideally, changes in upstream code should not have an effect on reproducing existing builds, until we want them to do so. A project's x.y version should always be built with the z.q dependency version if the software verification and testing have been done against this combination. The reason for this is that the smallest changes in an upstream dependency may affect the behavior of your software, even though there are no API or ABI changes. Your software will not have well-defined behavior if version consistency is not provided. Thus, having a way to provide version consistency is very crucial.

Ensuring version consistency in a super-build depends on the way that the super-build is organized. For repositories fetched over a version control system, it is relatively easy. Instead of cloning a project and using it as is, check out to a specific branch or tag. If there are no such anchor points, check out to a specific commit instead. This will future-proof your super-build. But even this may not be enough. Tags may get overridden, branches may be force-pushed, and history may be rewritten. To mitigate this risk, you may prefer to fork the project and use that as upstream instead. This way, you will have total control over the content that upstream has. But bear in mind that this method comes with a maintenance burden.

The moral of the story is, do not track an upstream blindly. Always keep an eye on recentchanges. For third-party dependencies that are consumed as archive files, always check for their hash digest. This way, you will ensure that you are indeed using the intended revision for your project, and if there is a change, you will have to manually acknowledge it.












































